---
title: "BST260 Final Report"
author: "Nick Birk"
date: "12/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Food security is a major public health concern. In order for people to live their healthiest lives, they must be able to access fresh produce. During my undergraduate studies, I worked part-time as a gardener and at a farmer’s market for two years, so food systems are of personal interest to me. During my time working these jobs, I learned about the way in which outside factors can largely influence an individual’s ability to access foods which are considered healthy. My hope is to observe what factors can predict the availability of various fruits and vegetables. Through this investigation, we may gain an understanding of what external forces are associated with the stability of our food systems and the health-related decisions that individuals are able to make.

One of the main priorities for this project is to create an interactive way to observe the trends of food availability over time, and how these metrics are associated with external factors. Such factors include real gross domestic product, diabetes prevalence, and unemployment rate. The ideal way to explore these trends would be through a Shiny app. In one tab, I will provide an interactive way to observe trends over time for a vegetable or fruit of choice, and in a second tab, I will provide an opportunity for the user to observe how availability correlates with other factors through scatterplots. Additionally, I hope to implement regression analyses to create a predictive system for food availability and consumption. 

# Data Assembly

We will begin by reading in the annual availability data, courtesty of the US Department of Agriculture. These data are given in the form of annual availabile pounds per capita for specific fruits and vegetables (10 of each are included here), and overall annual availability of fruits and vegetables per capita from 1970 to 2016.

```{r}
if(!require(tidyverse)) install.packages('tidyverse')
library(tidyverse)

#Total fruit and vegetable data
fruitveg <- read.csv('fruitveg.csv')
colnames(fruitveg)[1] <- 'year'

#Specific fruits and vegetables
apples <- read.csv('apples.csv')
avocados <- read.csv('avocados.csv')
bananas <- read.csv('bananas.csv')
bellpeppers <- read.csv('bellpeppers.csv')
broccoli <- read.csv('broccoli.csv')
cabbage <- read.csv('cabbage.csv')
carrots <- read.csv('carrots.csv')
eggplant <- read.csv('eggplant.csv')
grapes <- read.csv('grapes.csv')
lemons <- read.csv('lemons.csv')
mushrooms <- read.csv('mushrooms.csv')
onions <- read.csv('onions.csv')
oranges <- read.csv('oranges.csv')
peaches <- read.csv('peaches.csv')
pears <- read.csv('pears.csv')
pineapples <- read.csv('pineapples.csv')
potatoes <- read.csv('potatoes.csv')
spinach <- read.csv('spinach.csv')
tomatoes <- read.csv('tomatoes.csv')
watermelon <- read.csv('watermelon.csv')
```

Now, we can read in the data for potential covariates. Sources include the gapminder dataset, USDA, CDC, US Bureau of Labor Statistics, and Centers for Medicare & Medicaid Services. Note that diabetes prevlance data is not available for all years.

```{r}
covariates <- read.csv('outcomes.csv')
```

Note that farm income is in terms of 2019 dollars (i.e. adjusted for inflation).

```{r}
if(!require(dslabs)) install.packages('dslabs')
library(dslabs)

usgapminder <- gapminder %>% 
  filter(year %in% c(1970:2017) & country == 'United States') %>% 
  dplyr::select('year', 'life_expectancy', 'population', 'gdp')

covariates2 <- full_join(covariates, usgapminder, by = 'year')

#View(covariates2)
```

Now, we may combine the separate datasets. For purposes of linear regression, we want one form of the table in which every row represents a year. For purposes of visualization in the Shiny app, we want a second data frame in which every row represents a specific year-variable pairing.

```{r}
#First, we will join all tables on year
dat_annual <- full_join(fruitveg, covariates2, by = 'year')

head(dat_annual)
```

```{r}
#Now we create the dataset for our time series plots
dat_series <- inner_join(bananas, peaches, by = 'year')
colnames(dat_series)[2:3] <- c('bananas', 'peaches')
dat_series <- inner_join(dat_series, apples, by = 'year')
colnames(dat_series)[4] <- 'apples'
dat_series <- inner_join(dat_series, avocados, by = 'year')
colnames(dat_series)[5] <- 'avocados'
dat_series <- inner_join(dat_series, bellpeppers, by = 'year')
colnames(dat_series)[6] <- 'bellpeppers'
dat_series <- inner_join(dat_series, broccoli, by = 'year')
colnames(dat_series)[7] <- 'broccoli'
dat_series <- inner_join(dat_series, cabbage, by = 'year')
colnames(dat_series)[8] <- 'cabbage'
dat_series <- inner_join(dat_series, carrots, by = 'year')
colnames(dat_series)[9] <- 'carrots'
dat_series <- inner_join(dat_series, eggplant, by = 'year')
colnames(dat_series)[10] <- 'eggplant'
dat_series <- inner_join(dat_series, grapes, by = 'year')
colnames(dat_series)[11] <- 'grapes'
dat_series <- inner_join(dat_series, lemons, by = 'year')
colnames(dat_series)[12] <- 'lemons'
dat_series <- inner_join(dat_series, mushrooms, by = 'year')
colnames(dat_series)[13] <- 'mushrooms'
dat_series <- inner_join(dat_series, onions, by = 'year')
colnames(dat_series)[14] <- 'onions'
dat_series <- inner_join(dat_series, oranges, by = 'year')
colnames(dat_series)[15] <- 'oranges'
dat_series <- inner_join(dat_series, pears, by = 'year')
colnames(dat_series)[16] <- 'pears'
dat_series <- inner_join(dat_series, pineapples, by = 'year')
colnames(dat_series)[17] <- 'pineapples'
dat_series <- inner_join(dat_series, potatoes, by = 'year')
colnames(dat_series)[18] <- 'potatoes'
dat_series <- inner_join(dat_series, spinach, by = 'year')
colnames(dat_series)[19] <- 'spinach'
dat_series <- inner_join(dat_series, tomatoes, by = 'year')
colnames(dat_series)[20] <- 'tomatoes'
dat_series <- inner_join(dat_series, watermelon, by = 'year')
colnames(dat_series)[21] <- 'watermelon'


#UPDATE THIS WHENEVER YOU ADD SOMETHING
dat_series <- gather(dat_series, fruit, value, bananas:watermelon)

dat_series$emoji_code <- ''

dat_series$emoji_code[dat_series$fruit == 'bananas'] <- '1f34c'
dat_series$emoji_code[dat_series$fruit == 'peaches'] <- '1f351'
dat_series$emoji_code[dat_series$fruit == 'apples'] <- '1f34e'
dat_series$emoji_code[dat_series$fruit == 'avocados'] <- '1f951'
dat_series$emoji_code[dat_series$fruit == 'bellpeppers'] <- '1f336'
dat_series$emoji_code[dat_series$fruit == 'broccoli'] <- '1f966'
dat_series$emoji_code[dat_series$fruit == 'cabbage'] <-'1f957'
dat_series$emoji_code[dat_series$fruit == 'carrots'] <-'1f955'
dat_series$emoji_code[dat_series$fruit == 'eggplant'] <-'1f346'
dat_series$emoji_code[dat_series$fruit == 'grapes'] <-'1f347'
dat_series$emoji_code[dat_series$fruit == 'lemons'] <-'1f34b'
dat_series$emoji_code[dat_series$fruit == 'mushrooms'] <-'1f344'
dat_series$emoji_code[dat_series$fruit == 'onions'] <-'1f348'
dat_series$emoji_code[dat_series$fruit == 'oranges'] <-'1f34a'
dat_series$emoji_code[dat_series$fruit == 'pears'] <-'1f350'
dat_series$emoji_code[dat_series$fruit == 'pineapples'] <-'1f34d'
dat_series$emoji_code[dat_series$fruit == 'potatoes'] <-'1f954'
dat_series$emoji_code[dat_series$fruit == 'spinach'] <-'1f343'
dat_series$emoji_code[dat_series$fruit == 'tomatoes'] <-'1f345'
dat_series$emoji_code[dat_series$fruit == 'watermelon'] <-'1f349'
```


First, considering we wish to run a linear regression to predict produce availability, it would be helpful to check for potential correlation between all variables. In this way, we may avoid including multicollinear covariates.

```{r}
if(!require(corrplot)) install.packages('corrplot')
library(corrplot)
annualcor <-cor(dat_annual, use = 'pairwise.complete.obs')
corrplot(annualcor)
```

Indeed, it seems that diabetes prevalence, healthcare spending, life expectancy, population, and GDP have strong pairwise correlations with one another. We may want to consider a model which does not use all of these features, so as to keep our estimates more precise.

Further, we may use the included Shiny app to assess each pairwise correlation visually, as well as seeing the trend of each variable over time.

We will build and compare a few linear models.

Because of the temporality of the data, we want to be sure to include year as a predictor in all models. We will see how the regression model performs using year as the only predictor of total fruit and vegetable availability.

```{r}
lm.0 <- lm(total_fruit_veg ~ year, data = dat_annual)

summary(lm.0)
```

The model output shows that year is a statistically significant predictor of total fruit and vegetable availability. For each year, per capita total avaialbility increased on average by 1.9246 pounds. Let us now consider how the model performs when we adjust for each potential covariate from the dataset.

```{r}
lm.1 <- lm(total_fruit_veg ~ year + diabetes_percent + farm_income + unemployment + healthcare_percgdp + life_expectancy + population + gdp, data = dat_annual)

summary(lm.1)
```

We find that the model explains roughly 90% of the variability in total per capita fruit and vegetable availbility, as denoted by the adjusted R-squared value. This is quite impressive! However, as discussed above, many of the predictors were correlated with each other. Let us use a model which keeps only 2 of these correlated variables, diabetes prevalence and GDP.

```{r}
lm.2 <- lm(total_fruit_veg ~ year + diabetes_percent + farm_income + unemployment + gdp, data = dat_annual)

summary(lm.2)
```

Interestingly, the adjusted R-squared has only decreased slightly, changing from 0.9011 to 0.8767. Further, diabetes prevalence is now the only statistically significant predictor in the model. Let us now observe how the model performance changes if we remove GDP, leaving no predictors which are correlated with each other.

```{r}
lm.2 <- lm(total_fruit_veg ~ year + diabetes_percent + farm_income + unemployment, data = dat_annual)

summary(lm.2)
```

The adjusted R-squared has reduced even further to 0.8307, but all predictors aside from unemployment rate are now statistically significant. These results are quite interesting. In selecting the most appropriate model, we must decide which covariates necessarily must be adjusted for, even if they are correlated with each other. We must also remember that the sample size is quite small, so it will be important to investigate how these trends may or may not hold for larger datasets, such as ones which include data from other countries.

We will build a shiny app to allow customized visualizations as follows:

```{r}
if(!require(shiny)) install.packages('shiny')
library(shiny)
if(!require(shinythemes)) install.packages('shinythemes')
library(shinythemes)
```

Now we must get the emojis ready...
```{r}
devtools::install_github("dill/emoGG")
library(emoGG)
```

Got to do an experiment! How to change emojis :)
```{r, eval = FALSE}
froot <- '1f34c'

dat_series %>% filter(fruit == 'bananas') %>% ggplot(aes(x = year, y = value)) + geom_emoji(emoji = '1f348')
```

```{r}
cormat <- cor(dat_annual, use = 'pairwise.complete.obs')
cormat[8,]
```

We will now create a dataframe with cleaner variable names for use in the Shiny app.

```{r}
var_names <- c('Total Fresh Fruit Availability (lbs)' = 'fruit_fresh',
               'Total Processed Fruit Availability (lbs)' = 'fruit_processed',
               'Total Fresh Vegetable Availability (lbs)' = 'veg_fresh',
               'Total Processed Vegetable Availability (lbs)' = 'veg_processed',
               'Total Fruit Availability (lbs)' = 'total_fruit',
               'Total Vegetable Availability (lbs)' = 'total_veg',
               'Total Fruit and Vegetable Availability (lbs)' = 'total_fruit_veg',
               'Diabetes Prevalence (%)' = 'diabetes_percent',
               'Average Annual Farm Income ($)' = 'farm_income',
               'Annual Unemployment Rate (%)' = 'unemployment',
               'Percent GDP Spending on Healthcare' = 'healthcare_percgdp',
               'Average Life Expectancy (yrs)' = 'life_expectancy',
               'Population' = 'population',
               'GDP' = 'gdp')

var_names

names(var_names[which(var_names == 'life_expectancy')])
```

```{r}
veg_names <- c('Apples' = 'apples',
               'Avocados' = 'avocados',
               'Bananas' = 'bananas',
               'Bell Peppers' = 'bellpeppers',
               'Broccoli' = 'broccoli',
               'Cabbage' = 'cabbage',
               'Carrots' = 'carrots',
               'Eggplant' = 'eggplant',
               'Grapes' = 'grapes',
               'Lemons' = 'lemons',
               'Mushrooms' = 'mushrooms',
               'Onions' = 'onions',
               'Oranges' = 'oranges',
               'Peaches' = 'peaches',
               'Pears' = 'pears',
               'Pineapples' = 'pineapples',
               'Potatoes' = 'potatoes',
               'Spinach' = 'spinach',
               'Tomatoes' = 'tomatoes',
               'Watermelon' = 'watermelon')

veg_names

names(veg_names[which(veg_names == 'bellpeppers')])
```





```{r, eval = FALSE}
# Define UI
ui <- fluidPage( 
  theme = shinythemes::shinytheme("darkly"),
  titlePanel("US Per Capita Produce Availability"),
  
  tabsetPanel(
    
    
    tabPanel('Specific Products Time Series',
      sidebarLayout(
        sidebarPanel(
          selectInput("product", label = "Select a product",
                             choices = veg_names)),
      mainPanel(plotOutput('linePlot'))
               )),
    
    
    
    tabPanel('Variable Associations',
      fluidRow(
        column(4,
          selectInput('xvari', label = "Select a variable to plot on the x-axis", choices = var_names, selected = 'diabetes_percent'
                           ),
             selectInput('yvari', label = "Select a variable to plot on the y-axis", choices = var_names, selected = 'total_fruit_veg'
                           )),
             
      column(8, plotOutput("scatterPlot"))
    ),
    fluidRow(
      column(12, htmlOutput('texty'))
    ),
    fluidRow(
      column(6, plotOutput('scatterx')),
      column(6, plotOutput('scattery'))
    ))
  )
  )



# Define server
server <- function(input, output) {
     
  output$linePlot = renderPlot({
    if (input$product == 'bananas'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34c', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }
    
  else if (input$product == 'peaches'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f351', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }
    
  else if (input$product == 'apples'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34e', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    else if (input$product == 'avocados'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f951', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }  
    
     else if (input$product == 'bellpeppers'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f336', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    else if (input$product == 'broccoli'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f966', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
     else if (input$product == 'cabbage'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f957', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
     }  
    
     else if (input$product == 'carrots'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f955', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    else if (input$product == 'eggplant'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f346', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }  
    
    else if (input$product == 'grapes'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f347', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }  
    
     else if (input$product == 'lemons'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34b', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
     }  
    
        else if (input$product == 'mushrooms'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f344', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
        else if (input$product == 'onions'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f348', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
        else if (input$product == 'oranges'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34a', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
       else if (input$product == 'pears'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f350', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
       }  
    
        else if (input$product == 'pineapples'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34d', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
     else if (input$product == 'potatoes'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f954', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
     }  
    
         else if (input$product == 'spinach'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f343', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
         }  
    
             else if (input$product == 'tomatoes'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f345', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
             }  
    
                 else if (input$product == 'watermelon'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f349', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(names(veg_names[which(veg_names == input$product)]), ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    
   
  })
  
  output$texty <- renderUI({
    HTML('Trends for each variable over time are included below:')
  })
  
  output$scatterPlot = renderPlot({
    dat_annual %>% ggplot(aes_string(x = input$xvari, y = input$yvari)) + geom_point() + ggtitle(paste0(names(var_names[which(var_names == input$yvari)]), ' by ', names(var_names[which(var_names == input$xvari)]))) + xlab(names(var_names[which(var_names == input$xvari)])) + ylab(names(var_names[which(var_names == input$yvari)]))  + geom_smooth()
  })
  
  output$scatterx = renderPlot({
    dat_annual %>% ggplot(aes_string(x = 'year', y = input$xvari)) + geom_point() + ggtitle(paste0(names(var_names[which(var_names == input$xvari)]), ' by year')) + xlab('Year') + ylab(names(var_names[which(var_names == input$xvari)]))  + geom_smooth()
  })
  
  output$scattery = renderPlot({
    dat_annual %>% ggplot(aes_string(x = 'year', y = input$yvari)) + geom_point() + ggtitle(paste0(names(var_names[which(var_names == input$yvari)]), ' by year')) + xlab('Year') + ylab(names(var_names[which(var_names == input$yvari)]))  + geom_smooth()
  })
  

}

shinyApp(ui = ui, server = server)
```










And now, APP WITH EMOJIS :)

```{r, eval = FALSE}
# Define UI
ui <- fluidPage( 
  theme = shinythemes::shinytheme("darkly"),
  titlePanel("US Per Capita Produce Availability"),
  
  tabsetPanel(
    
    
    tabPanel('Specific Products Time Series',
      sidebarLayout(
        sidebarPanel(
          selectInput("product", label = "Select a product",
                             choices = sort(unique(dat_series$fruit)))),
      mainPanel(plotOutput('linePlot'))
               )),
    
    
    
    tabPanel('Variable Associations',
      fluidRow(
        column(4,
          selectInput('xvari', label = "Select a variable to plot on the x-axis", choices = var_names, selected = 'diabetes_percent'
                           ),
             selectInput('yvari', label = "Select a variable to plot on the y-axis", choices = var_names, selected = 'total_fruit_veg'
                           )),
             
      column(8, plotOutput("scatterPlot"))
    ),
    fluidRow(
      column(12, htmlOutput('texty'))
    ),
    fluidRow(
      column(6, plotOutput('scatterx')),
      column(6, plotOutput('scattery'))
    ))
  )
  )



# Define server
server <- function(input, output) {
     
  output$linePlot = renderPlot({
    if (input$product == 'bananas'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34c', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }
    
  else if (input$product == 'peaches'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f351', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }
    
  else if (input$product == 'apples'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34e', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    else if (input$product == 'avocados'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f951', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }  
    
     else if (input$product == 'bellpeppers'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f336', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    else if (input$product == 'broccoli'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f966', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
     else if (input$product == 'cabbage'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f957', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
     }  
    
     else if (input$product == 'carrots'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f955', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    else if (input$product == 'eggplant'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f346', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }  
    
    else if (input$product == 'grapes'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f347', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
    }  
    
     else if (input$product == 'lemons'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34b', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
     }  
    
        else if (input$product == 'mushrooms'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f344', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
        else if (input$product == 'onions'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f348', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
        else if (input$product == 'oranges'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34a', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
       else if (input$product == 'pears'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f350', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
       }  
    
        else if (input$product == 'pineapples'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f34d', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
        }  
    
     else if (input$product == 'potatoes'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f954', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
     }  
    
         else if (input$product == 'spinach'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f343', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
         }  
    
             else if (input$product == 'tomatoes'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f345', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0(input$product, ' availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
             }  
    
                 else if (input$product == 'watermelon'){
    dat_series %>% 
      filter(fruit == input$product) %>%
      ggplot(aes(x = year, y = value)) + 
      geom_emoji(emoji = '1f349', size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle(paste0('Watermelon availbility over time')) + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  }  
    
    
   
  })
  
  output$texty <- renderUI({
    HTML('Trends for each variable over time are included below:')
  })
  
  output$scatterPlot = renderPlot({
    dat_annual %>% ggplot(aes_string(x = input$xvari, y = input$yvari)) + geom_point() + ggtitle(paste0(names(var_names[which(var_names == input$yvari)]), ' by ', names(var_names[which(var_names == input$xvari)]))) + xlab(names(var_names[which(var_names == input$xvari)])) + ylab(names(var_names[which(var_names == input$yvari)]))  + geom_smooth()
  })
  
  output$scatterx = renderPlot({
    dat_annual %>% ggplot(aes_string(x = 'year', y = input$xvari)) + geom_point() + ggtitle(paste0(names(var_names[which(var_names == input$xvari)]), ' by year')) + xlab('Year') + ylab(names(var_names[which(var_names == input$xvari)]))  + geom_smooth()
  })
  
  output$scattery = renderPlot({
    dat_annual %>% ggplot(aes_string(x = 'year', y = input$yvari)) + geom_point() + ggtitle(paste0(names(var_names[which(var_names == input$yvari)]), ' by year')) + xlab('Year') + ylab(names(var_names[which(var_names == input$yvari)]))  + geom_smooth()
  })
  

}

shinyApp(ui = ui, server = server)
```




