---
title: "Full Report"
author: "Nick Birk"
date: "11/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We will investigate annual fruit and vegetable availability in America.

# Data Assembly

We will begin by reading in the annual availability data, courtesty of the US Department of Agriculture. These data are given in the form of annual availabile pounds per capita for specific fruits and vegetables (10 of each are included here), and overall annual availability of fruits and vegetables per capita from 1970 to 2016.

```{r}
if(!require(tidyverse)) install.packages('tidyverse')
library(tidyverse)

#Total fruit and vegetable data
fruitveg <- read.csv('fruitveg.csv')
colnames(fruitveg)[1] <- 'year'

#Specific fruits and vegetables
apples <- read.csv('apples.csv')
avocados <- read.csv('avocados.csv')
bananas <- read.csv('bananas.csv')
bellpeppers <- read.csv('bellpeppers.csv')
broccoli <- read.csv('broccoli.csv')
cabbage <- read.csv('cabbage.csv')
carrots <- read.csv('carrots.csv')
eggplant <- read.csv('eggplant.csv')
grapes <- read.csv('grapes.csv')
lemons <- read.csv('lemons.csv')
mushrooms <- read.csv('mushrooms.csv')
onions <- read.csv('onions.csv')
oranges <- read.csv('oranges.csv')
peaches <- read.csv('peaches.csv')
pears <- read.csv('pears.csv')
pineapples <- read.csv('pineapples.csv')
potatoes <- read.csv('potatoes.csv')
spinach <- read.csv('spinach.csv')
tomatoes <- read.csv('tomatoes.csv')
watermelon <- read.csv('watermelon.csv')
```

Now, we can read in the data for potential covariates. Sources include the gapminder dataset, USDA, CDC, US Bureau of Labor Statistics, and Centers for Medicare & Medicaid Services. Note that diabetes prevlance data is not available for all years.

```{r}
covariates <- read.csv('outcomes.csv')
```

Note that farm income is in terms of 2019 dollars (ie adjusted for inflation).

```{r}
if(!require(dslabs)) install.packages('dslabs')
library(dslabs)

usgapminder <- gapminder %>% 
  filter(year %in% c(1970:2017) & country == 'United States') %>% 
  dplyr::select('year', 'life_expectancy', 'population', 'gdp')

covariates2 <- full_join(covariates, usgapminder, by = 'year')

#View(covariates2)
```

Now, we may combine the separate datasets. For purposes of linear regression, we want one form of the table in which every row represents a year. For purposes of visualization in the Shiny app, we want a second data frame in which every row represents a specific year-variable pairing.

```{r}
#First, we will join all tables on year
dat_annual <- full_join(fruitveg, covariates2, by = 'year')

head(dat_annual)
#dat_by_var 
```


First, considering we wish to run a linear regression to predict produce availability, it would be helpful to check for potential correlation between all variables. In this way, we may avoid including multicollinear covariates.

```{r}
if(!require(corrplot)) install.packages('corrplot')
library(corrplot)
annualcor <-cor(dat_annual, use = 'pairwise.complete.obs')
corrplot(annualcor)
```

Indeed, it seems that diabetes prevalence, healthcare spending, life expectancy, population, and GDP have strong pairwise correlations with one another. We may want to consider a model which does not use all of these features, so as to keep our estimates more precise.



We will build a shiny app to allow customized visualizations as follows:

```{r}
if(!require(shiny)) install.packages('shiny')
library(shiny)
if(!require(shinythemes)) install.packages('shinythemes')
library(shinythemes)



```

Now we must get the emojis ready...
```{r}
devtools::install_github("dill/emoGG")
library(emoGG)
fruitlist = c('bananas', 'peaches')
```


Here is a sample app:
```{r, eval = FALSE}
# Define UI
ui <- fluidPage( 
  theme = shinythemes::shinytheme("darkly"),
  titlePanel("US Per Capita Produce Availability"),
  
  tabsetPanel(
    
    
    tabPanel('Specific Products Time Series',
      sidebarLayout(
        sidebarPanel(
          selectInput("product", label = "Select a product",
                             choices = fruitlist)),
      mainPanel(plotOutput('linePlot'))
               )),
    
    
    
    tabPanel('Variable Associations',
      sidebarLayout(
        sidebarPanel(
          selectInput('xvari', label = "Select a variable to plot on the x-axis", choices = colnames(dat_annual)[2:15], selected = 'diabetes_percent'
                           ),
             selectInput('yvari', label = "Select a variable to plot on the y-axis", choices = colnames(dat_annual)[2:15], selected = 'total_fruit_veg'
                           )),
             
      mainPanel(plotOutput("scatterPlot"))
    ))
  )
  )



# Define server
server <- function(input, output) {
  
    gap_year <- reactive(gapminder %>% filter(year == input$year))
    
  output$linePlot = renderPlot({
    bananas %>% 
      ggplot(aes(x = year, y = total)) + 
      geom_emoji(emoji="1f34c", size = 0.04) + 
      geom_line(alpha = 0.3) + ggtitle('Banans!') + xlab('Year') +
      ylab('Per capita availability (lbs)') 
  })
  
  output$scatterPlot = renderPlot({
    dat_annual %>% ggplot(aes_string(x = input$xvari, y = input$yvari)) + geom_point() + ggtitle(paste0(input$yvari, ' by ', input$xvari)) + xlab(input$xvari) + ylab(input$yvari)  + geom_smooth()
  })
  
  

}

shinyApp(ui = ui, server = server)
```


